---
title: "Population stability analysis for Balanus improvisus"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library("rprojroot")
root_crit <- has_dirname("PWS_2_5_west_coast_model", subdir = "src")
root_dir <- root_crit$make_fix_file()

library("tidyr")
library("dplyr")
library("ggplot2")
library("popbio")
library("fs")

suppressPackageStartupMessages(library("DEoptim"))
options(scipen = -4)
```

Functions to determine proportion of overall survival that is transferred between
classes (gamma), and the probability of surviving and staying in the same class is
$P_i$ and the probability of surviving and transferring between classes is $G_i$
 (From Caswell 2001).

```{r}
gamma_i <- function(sigma_i, lam, T_i){
  res <- ((sigma_i / lam) ^ T_i - (sigma_i / lam) ^ (T_i - 1)) /
    ((sigma_i / lam) ^ T_i - 1)
}


P_i <- function(p, T_i){
  p * (1 - p ^ (T_i - 1) ) / (1 - p ^ T_i)
}

G_i <- function(p, T_i) {
  p ^ T_i * (1 - p) / (1 - p ^ T_i)
}

```

The overall survival for each stage (sigma), and the duration of each stage
(T) in units of 6-hour timesteps.
```{r survivalprob}
balanus_lh <- list("larva" = list("sigma" = 0.515, "T" = 44),
                   "cyprid" = list("sigma" = 0.94, "T" = 12),
                   "juvenile" = list("sigma" = 0.58, "T" = 84),
                   "adult" = list("sigma" = 0.998, "T" = 2192))
```


The function used to estimate the value for $\gamma_{juvenile}$ for an overall
dominant eigenvalue of 1.0006.

```{r fit-gamma}
balanus_fit <- function(log_gamma_3) {
  lam <- 1.0006

  sigma_1 <- balanus_lh[["larva"]][["sigma"]]
  gamma_1 <- 1/balanus_lh[["larva"]][["T"]]
  P_1 <- sigma_1 * (1 - gamma_1)
  G_1 <- sigma_1 * gamma_1

  sigma_2 <- balanus_lh[["cyprid"]][["sigma"]]
  gamma_2 <- gamma_i(sigma_i = sigma_2, lam = lam,
    T_i = balanus_lh[["cyprid"]][["T"]])

  P_2 <- sigma_2 * (1 - gamma_2)
  G_2 <- sigma_2 * gamma_2

  sigma_3 <- balanus_lh[["juvenile"]][["sigma"]]

  gamma_3 <- exp(log_gamma_3)
  P_3 <- sigma_3 * (1 - gamma_3)
  G_3 <- sigma_3 * gamma_3

  sigma_4 <- balanus_lh[["adult"]][["sigma"]]

  gamma_4 <- gamma_i(sigma_i = sigma_4, lam = lam,
  	T_i =  balanus_lh[["adult"]][["T"]])

  P_4 <- sigma_4 * (1 - gamma_4)

  F_4 <- 13.69 # 10000 eggs per season

  A <- matrix(c(P_1, 0, 0, F_4,
                G_1, P_2, 0, 0,
                0, G_2, P_3, 0,
                0, 0, G_3, P_4), nrow = 4, byrow = TRUE)
  gammas <- c("gamma1" = gamma_1, "gamma2" = gamma_2, "gamma3" = gamma_3,
    "gamma4" = gamma_4)
  res = list(A = A, gammas = gammas)
}
```


```{r estimate-juv-gamma}
gamma_min_function <- function(gamma) {
  A_local <- balanus_fit(gamma)[["A"]]
  A_local[is.nan(A_local)] <- .Machine$double.eps

  res <- (1.0006 - eigen.analysis(A_local)$lambda1)^2
}

```

The function to minimize.
```{r minFunction}

gamma_min_function <- function(gamma) {
  A_local <- balanus_fit(gamma)[["A"]]
  A_local[is.nan(A_local)] <- .Machine$double.eps

  res <- (1.0006 - eigen.analysis(A_local)$lambda1)^2
}


res <- DEoptim(gamma_min_function, lower = -7, upper = 0,
  control = list(trace = 0))
```

The resulting population transition matrix is

```{r, echo=FALSE, results='asis'}
A_fit <- balanus_fit(res$optim$bestmem)[["A"]]
rownames(A_fit) <- c("larva", "cyprid", "juvenile", "adult")
colnames(A_fit) <- c("larva", "cyprid", "juvenile", "adult")
knitr::kable(A_fit)
```

with $\gamma$ of each stage as:

```{r gamma-output, echo = FALSE, results='asis'}
A_gammas <- balanus_fit(res$optim$bestmem)[["gammas"]]
names(A_gammas) <- c("larva", "cyprid", "juvenile", "adult")
knitr::kable(A_gammas)
```

```{r run-model, echo = FALSE}
source(path(root_dir(), "munge", "01-model_control_pop_stability.R"))

```
After submitting the population transition matrix to a model where there is
no ship movement, the average population size for each lifestage in the seed
ports is shown as:

```{r plot-log, echo=FALSE}
pop_stability_results <- readr::read_log(file.path(root_dir(), "logs",
  "parameter002_stab_ports_pop_trace.log"), col_names = FALSE)

names(pop_stability_results) <- c("log_level", "datetime", "parameter",
  "variable", "time_idx", "larva", "cyprid", "juvenile", "adult")

port_stability <- pop_stability_results %>%
  select(time_idx, larva, cyprid, juvenile, adult) %>%
  pivot_longer(., cols = -time_idx, names_to = "lifestage", values_to = "population") %>%
  mutate(lifestage = factor(lifestage, levels = c("larva", "cyprid", "juvenile",
    "adult")))

ggplot(port_stability, aes(time_idx, population)) +
  facet_wrap(~lifestage, ncol = 1, nrow = 4, scales = "free_y") +
  geom_line() +
  labs(x = "time index", y = "population") +
  theme_bw()

```
